---
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    theme: paper
    css: ["C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/template/css/bootstrap-theme.css","C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/template/css/bootstrap.css"]
  pdf_document: default
  word_document:
    reference_docx: ../template/word-styles-reference-01.docx
editor_options: 
  chunk_output_type: console
params:
  species: 
    value: x
  genus:
    value: x
  subgenus:
    value: 
  s_rank:
    value: 
  common_name:
    value: "no common name"
  account:
    value: ""
  locationUncertainty:
    value: 250
---
<style>
.center {
  text-align: center;
}


</style>


<div class="row">
<div class="col-xs-12 center-block text-center">
<p style="font-size: 32pt; text-align: center; width: 100%;"><strong>Vermont's Encyclopedia of Life</strong></p>
</div>
</div>

```{css, echo = FALSE, eval = TRUE}
/* see https://github.com/rstudio/rmarkdown/issues/1949 */

.leaflet-layer  {
    min-width: 400px;
}
.leaflet-tile-container {
    width: 100%;
}
.leaflet-container { 
    background: #FFFFFF; 
}
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# READ IN PACKAGES AND GET DATA FROM GBIF
source("C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/SpeciesAccount_childs/01_species_account_load_packages_data_from_GBIF.R");

NumiNatRecords <- length(grep(pattern = "inaturalist", x = samp_spp$occurrenceID))
```

```{r echo = FALSE, warnings = FALSE}
# Gather phenology info
source("C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/SpeciesAccount_childs/02_species_account_phenology.R")
```

```{r, echo = FALSE, message = FALSE, results = 'hide', error = FALSE, warning = FALSE}
source("C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/SpeciesAccount_childs/03_species_account_iNat_photos.R")
```

```{r echo = FALSE, message = FALSE, error = FALSE, warning = FALSE, results='hide'}
source("C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/SpeciesAccount_childs/04_species_account_AssociatedTaxa.R")
```

<div class="row">
 <div class="col-xs-6">
 <p style="text-align: center; font-size: 32pt;"> <i>`r params$species`</i> </p>
 <p style="text-align: left; font-size: 16pt;"><strong>Provisional S-rank:</strong> `r params$s_rank`</p>  
 <p style="text-align: left; font-size: 16pt;"><strong>Genus:</strong> `r params$genus` `r params$subgenus`</p>    
  
<p style="text-align: left; font-size: 16pt;"><strong>Common name:</strong> `r params$common_name`</p>
<p style="text-align: left; font-size: 14pt;"><strong>First VT record:</strong> `r min(samp_spp$year,na.rm = TRUE)`</p>
<p style="text-align: left; font-size: 14pt;"><strong>Last VT record:</strong> `r max(samp_spp$year,na.rm = TRUE)`</p>
<p style="text-align: left; font-size: 14pt;"><strong>VT records:</strong> `r prettyNum(nrow(samp_spp[!is.na(samp_spp$gbifID),]),big.mark = ",")` </p> 
<p style="text-align: left; font-size: 14pt;"><strong>iNaturalist records:</strong> `r prettyNum(NumiNatRecords,big.mark = ",")`</p>
</div>

<div class = "col-xs-6">
<img style="padding: 3px 3px 0px 3px;" src=`r imgURL` alt=`r iNatPhotos$results$taxon$species`></img>
<p style="text-align: center; font-size: 12pt;"> `r imageCredit`</p>
 </div>
</div>

<div class = "row">
<div class="container" style="border-radius: 10px; background-color: lightgray; padding: 5px 5px 5px 5px; border: 1px solid; box-shadow: 5px 5px #888888;">
<p style="text-align: left; font-size: 14pt;"> `r params$account` </p>
</div>
</div>

<div class="row">
<div class = "col-xs-5">
<p style="font-size: 16pt; text-align: center;"><strong>Phenology</strong></p><i class="fa-solid fa-circle-info"></i>
 
```{r, echo = FALSE, dpi = 400, warning = FALSE, comment = NA}
source("C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/SpeciesAccount_childs/05_species_account_Plot_phenology.R")
```
</div>

<div class = "col-xs-7">
<p style="font-size: 16pt; text-align: center;"><strong>Flower interactions</strong></p>   
```{r message = FALSE, error = FALSE, warning = FALSE, echo = FALSE, comment = NA}
if(nrow(AssociatedTaxa)!=0){
library(kableExtra)
kableExtra::kbl(BeePropData[,c("Vegetation","Observations","Proportion")], row.names = FALSE,
                center = FALSE,
                caption = "",
                align = c("l","c","c"),
                col.names = c("Vegetation",
                              "Observations",
                              "Proportion"),
                format = "html",
                midrule = "c",
                escape = FALSE) %>%
     kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                   font_size = 18) %>%
    kableExtra::scroll_box(height = "350px")}else{
 cat("<p> No interaction data are available for this species</p>")
}
```

</div>
</div>

<div class = "row h-50">
<div class="col-xs-12 center-block text-center">
<p style="font-size: 16pt; text-align: center;"><strong>Vermont observations</strong></p>

```{r, echo = FALSE, error = FALSE, message = FALSE, warning = FALSE,out.width = '95%'}
source("C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/SpeciesAccount_childs/06_species_account_Plot_yearlyrecords.R")
```

</div> 
</div>

<div class = "row h-100">
<div class="col-xs-12">
<p style="font-size: 16pt; text-align: center;"><strong>Distribution</strong></p>

```{r eval = FALSE, echo = FALSE, message = FALSE, dpi=400, out.height="90%"}
#THIS CHUNK CREATES A STATIC BASIC MAP

source("C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/SpeciesAccount_childs/07_species_account_Plot_vt_locations_towns.R")
```

```{r eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE}

# THIS CHUNK CREATES A MOUSEOVER OF THE OBSERVATIONS
source("C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/SpeciesAccount_childs/07a_species_account_Plot_vt_locations_towns_mouseover.R")
```

```{r eval = TRUE, echo = FALSE, warning = FALSE, out.height = "99%"}
leafletPlot
```

</div>
</div> 

<div class = "row">
<div class = "col-xs-6">
<p style="font-size: 16pt; text-align: center;"><strong>Habitat</strong></p> 

```{r, echo = FALSE, dpi = 400}
source("C:/Users/mhall/State_Of_Biodiversity/Bee_Monograph_Hardy_et_al/SpeciesAccount_childs/08_species_account_Plot_habitat_observations.R")
```
</div>

<div class = "col-xs-6">
<p style="font-size: 16pt; text-align: center;"><strong>Elevation</strong></p> 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Query the database to get habitat of observations # 

VTelev <- terra::rast("C:/Users/mhall/State_Of_Biodiversity/spatial_data/srtm_merged.tif")
VTterra <- terra::vect(as(VTshape,"Spatial"))
VTelevs <- terra::extract(VTelev, VTterra)

vt_xy <- density(VTelevs$srtm_merged, bw = 0.25)

elevOcc <- dbGetQuery(GBIFDb,
                  paste0('SELECT "elevation_m" as elev
                          FROM occurrence LEFT OUTER JOIN obs_environmental ON (occurrence."species" = obs_environmental."species")
                          WHERE occurrence."species" = \'',params$species,'\'
                          AND occurrence."coordinateUncertaintyInMeters" <', params$locationUncertainty))
                          
# summarize elevation 
if(length(elevOcc$elev) >= 5){
elev_xy <- density(as.numeric(elevOcc$elev))

par(bty ="n", mar = c(1,3.8,1,1))
plot(NA, xlim = c(0,5), ylim = c(0,1400),
     las = 1, 
     xaxt = "n",
     ylab = "Elevation (m)",
     xlab = "")

polygon(x = vt_xy$y*100,
        y = vt_xy$x,
        col = "gray88",
        border = "gray88")

polygon(x = elev_xy$y*100,
        y = elev_xy$x,
        col = "gray50",
        border = "gray50")

points(x=0, y = median(as.numeric(elevOcc$elev),na.rm = TRUE), pch = 19, cex = 1.25)
segments(x0 = 0, x1 = 0, 
         y0 = quantile(as.numeric(elevOcc$elev),0.25),
         y1 = quantile(as.numeric(elevOcc$elev),0.75))
}else{
  cat("Too few observations to create elevational distribution")
}

```
</div>
</div>

<div class = "row">
<div class = "col-xs-12 mx-auto">
<p style="font-size: 16pt; text-align: center;"><strong>Sampling</strong></p> 

```{r, eval = TRUE, echo = FALSE, message = FALSE, error = FALSE, results = 'hide', dpi = 400}

# Sampling design # 
library(googlesheets4)
library(rmarkdown)
library(jsonlite)
library(rgbif)
library(sf)
library(raster)
library(DBI)
library(RPostgres)
library(rpostgis)
library(dbplyr)
library(dplyr)
library(rgdal)
library(dismo)
library(ENMeval)
library(doSNOW)
library(spatstat)
library('Ternary')


GBIFDb <- DBI::dbConnect(RPostgres::Postgres(),
                         dbname = 'vt-gbif-data',
                         host = 'localhost',
                         port = 5432, # or any other port specified by your DBA
                         user = "postgres",
                         password = "password")

# link to the google sheet
google_sheet <- "https://docs.google.com/spreadsheets/d/1MfeCwgbFn2-twocqi3db-2LXreHsBHcABuRtvR6qL0Q/edit#gid=1494373398"

# get the properties of the google sheet
gs_properties <- sheet_properties(google_sheet)

# read in the sheet of interest 
protocolSampClass <- range_read(google_sheet, sheet = which(gs_properties$name == "Sampling_protocols_types"))

# Grab the data from our local dataset # 
bees <- dbGetQuery(GBIFDb,
                   'SELECT "gbifID","species","speciesKey","genus", "samplingProtocol","eventDate"
                    FROM occurrence 
                    WHERE "familyKey" = 4334 OR
                          "familyKey" = 7901 OR
                          "familyKey" = 7908 OR
                          "familyKey" = 4345 OR
                          "familyKey" = 7905 OR
                          "familyKey" = 7911;')

# new column for the re-structured protocols 
bees$newSampProtocol <- NA

# how many are there 
nProto <- length(unique(protocolSampClass$samplingProtocal))

# use the google sheet to rename them
for(i in 1:nProto){
bees$newSampProtocol[which(bees$samplingProtocol == protocolSampClass$samplingProtocal[i])] <- protocolSampClass$Method[i]
}

# Summarize the new protocol for each species # 
bsamp <- tapply(bees$gbifID,
       list(bees$species,
            bees$newSampProtocol),
       FUN = function(x){length(unique(x))})


# keep only the classes Spencer wants # 
bsampling <- bsamp[,c("Bowl","Malaise Trap","Net")]

# convert to percentage 
bSampPercent <- (bsampling/apply(bsampling,1,sum,na.rm = TRUE))*100

# convert NA to zeros 
bSampPercent[is.na(bSampPercent)] <- 0

# genus color 
# genus 


genus <- unique(bees$genus)

genusCol <- viridis::viridis(n = length(genus),option = "H",alpha = 0.35)

gcolor <- genusCol[match(gsub( " .*$", "",row.names(bSampPercent) ),genus)]

bSampPerDF <- data.frame(bSampPercent,gcolor)

par(mar = c(1,1,1,1), xpd = TRUE)
TernaryPlot(alab = "Bowl", 
            blab = "Malaise Trap", 
            clab = "Net",
            point = 'up', 
            lab.cex = 0.8, 
            grid.minor.lines = 0,
            grid.lty = 'solid', 
            col = rgb(0.9, 0.9, 0.9),
            grid.col = 'gray88', 
            ticks.length = 0.02,
            axis.cex = 0.8,
            axis.col = rgb(0.6, 0.6, 0.6), 
            ticks.col = rgb(0.6, 0.6, 0.6),
            axis.rotate = FALSE)

apply(bSampPerDF,
      1,function(x){
        if(all(as.numeric(x[1:3])==0)){
          TernaryPoints(cbind(33,33,33),pch = 19, 
                        col = rgb(255,255,255,100,max = 255))
        }else{
        TernaryPoints(as.numeric(x[1:3]), 
                      pch = 19, 
                      col = rgb(150,150,150,100, max = 255))}
        })

sppXYZ <- bSampPerDF[which(row.names(bSampPerDF) == params$species),]

if(!all(sppXYZ[1:3]==0)){
TernaryPoints(sppXYZ[1:3],
              pch = 19,
              col = "black",
              cex = 1.25)} 
```
</div>
</div>
